---
title: "Project setup"
author: "jennysjaarda"
date: "2021-06-14"
output: workflowr::wflow_html
editor_options:
    toc: true
    number_sections: true
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

**To reproduce the results from this project, please follow these instructions.**

*In general, [`targets`](https://books.ropensci.org/targets/dynamic.html) was used to manage long-running code and [`workflowr`](https://jdblischak.github.io/workflowr/index.html) was used to manage the [website](https://jennysjaarda.github.io/proxymr/index.html).*

***


# Initiate project on remote server.

All processing scripts were run from the root sgg directory.
Project was initialized using `workflowr` rpackage, see [here](https://jdblischak.GitHub.io/workflowr/articles/wflow-01-getting-started.html).

On personal computer:

```{r eval=FALSE}
project_name <- "proxyMR"
library("workflowr")

wflow_start(project_name) # creates directory called `project_name`

options("workflowr.view" = FALSE) # if using cluster
options("workflowr.view" = TRUE) # defaule, if using Rstudio on personal computer

wflow_build() # create directories
options(workflowr.sysgit = "")

wflow_use_github("jennysjaarda") # select option 2: manually create new repository

wflow_git_push()
```

You have now successfully created a GitHub repository for your project that is accessible on GitHub and the servers.

Next setup a local copy.


# Create local copy on personal computer.

Within terminal of personal computer, clone the git repository.

```{r, engine = 'bash', eval = FALSE}
cd ~/Dropbox/UNIL/projects/
git clone https://GitHub.com/jennysjaarda/proxyMR.git proxyMR
```

Open project in atom (or preferred text editor) and modify the following files:

- Because Terminal cannot generate a preview and `workflowr` doesn't like the sysgit, to the `.Rprofile` file, add:
  - `options(workflowr.sysgit = "")`
  - `options("workflowr.view" = FALSE)`
- To ensure git hub isn't manaaging large files, modify the `.gitignore` file, by adding the following lines:
  - `analysis/*`
  - `data/*`
  - `!analysis/*.Rmd`
  - `!data/*.md`
  - `.git/`
  - `.drake/`
- Save and push these changes to GitHub.
- Pull to the server.



